Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-04-06 paper.qmd
Stash

```{r}
#| label: umap-pbmc-read-best
umap_pbmc2 <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
pbmc_scaled_obj <- gen_scaled_data(
  data = umap_pbmc2)
umap_pbmc_scaled_best <- pbmc_scaled_obj$scaled_nldr

@@

<!--Fit the best model and compute error-->
```{r}
#| label: num-bins-pbmc

@@

algo_obj_pbmc <- gen_nldr_vis_algo_obj(
  high_d_data = training_data_pbmc, 
  nldr_data = umap_pbmc2, 
  num_x_bins = num_bins_x_pbmc)

umap_pbmc_scaled_best <- algo_obj_pbmc$nldr_scaled
distance_pbmc <- algo_obj_pbmc$distance_df
tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
benchmark_pbmc <- algo_obj_pbmc$benchmark
df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
df_bin_pbmc <- algo_obj_pbmc$df_bin

@@

umap_pbmc_scaled_best_with_cluster <- umap_pbmc_scaled_best |>
  mutate(cluster = umap_pbmc_scaled_with_cluster$cluster)

trimesh_removed_pbmc_best <- ggplot() + 
  geom_point(
    data = umap_pbmc_scaled_best_with_cluster,
    aes(
      x = tSNE1,
      y = tSNE2,
      color = cluster
    ),
    alpha = 0.3#,
    #color = "#999999"
  ) +
    geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
  interior_annotation("b1", cex = 2) +
  theme(
    aspect.ratio = 1
  )
```

```{r}
#| label: prep-pbmc-umap-model-proj-best

@@

```{r}
#| label: langevitour-pbmc-umap-best-proj
#| eval: false

@@

```{r}
#| label: pbmc-umap-model-proj-best

@@

pbmc_proj_umap_model1_best <- plot_proj(
    proj_obj = proj_obj1,
    point_param = c(1.5, 0.2), # size, alpha, color
    plot_limits = c(-0.65, 0.65), 
    title = "b2", 
    cex = 2, 
    axis_text_size = 5,
    is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))

@@

pbmc_proj_umap_model2_best <- plot_proj(
    proj_obj = proj_obj2,
    point_param = c(1.5, 0.2), # size, alpha, color
    plot_limits = c(-1, 0.75), 
    title = "b3", 
    cex = 2, 
    axis_text_size = 5,
    is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
```

@@

free(trimesh_removed_pbmc) + free(trimesh_removed_pbmc_best) +
  pbmc_proj_umap_model1 + pbmc_proj_umap_model1_best +
  pbmc_proj_umap_model2 + pbmc_proj_umap_model2_best +
  plot_layout(nrow=2, byrow=FALSE) &
  theme(legend.position='none')
```

>>>>

```{r}
#| label: tsne-pbmc-read-best
tsne_pbmc <- read_rds("data/pbmc3k/pbmc_tsne_30.rds")
pbmc_scaled_obj <- gen_scaled_data(
  data = tsne_pbmc)
tsne_pbmc_scaled_best <- pbmc_scaled_obj$scaled_nldr

@@

<!--Fit the best model and compute error-->
```{r}
#| label: tsne-num-bins-pbmc

@@

algo_obj_pbmc <- gen_nldr_vis_algo_obj(
  high_d_data = training_data_pbmc, 
  nldr_data = tsne_pbmc, 
  num_x_bins = num_bins_x_pbmc)

tsne_pbmc_scaled_best <- algo_obj_pbmc$nldr_scaled
distance_pbmc <- algo_obj_pbmc$distance_df
tr_from_to_df_pbmc <- algo_obj_pbmc$tr_from_to_df
benchmark_pbmc <- algo_obj_pbmc$benchmark
df_bin_centroids_pbmc <- algo_obj_pbmc$df_bin_centroids
df_bin_pbmc <- algo_obj_pbmc$df_bin

@@

tsne_pbmc_scaled_best_with_cluster <- tsne_pbmc_scaled_best |>
  mutate(cluster = umap_pbmc_scaled_with_cluster$cluster)

trimesh_removed_pbmc_best <- ggplot() + 
  geom_point(
    data = tsne_pbmc_scaled_best_with_cluster,
    aes(
      x = tSNE1,
      y = tSNE2,
      color = cluster
    ),
    alpha = 0.3#,
    #color = "#999999"
  ) +
  geom_segment(data = tr_from_to_df_pbmc, 
               aes(
                 x = x_from, 
                 y = y_from, 
                 xend = x_to, 
                 yend = y_to),
               colour = "#000000",
               linewidth = 1) +
  scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada")) +
  interior_annotation("b1", cex = 2) +
  theme(
    aspect.ratio = 1
  )
```

```{r}
#| label: prep-pbmc-tsne-model-proj-best

@@

```{r}
#| label: langevitour-pbmc-tsne-best-proj
#| eval: false

@@

```{r}
#| label: pbmc-tsne-model-proj-best

@@

pbmc_proj_tsne_model1_best <- plot_proj(
  proj_obj = proj_obj1,
  point_param = c(1.5, 0.2), # size, alpha, color
  plot_limits = c(-0.65, 0.65), 
  title = "b2", 
  cex = 2, 
  axis_text_size = 5,
  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))

@@

pbmc_proj_tsne_model2_best <- plot_proj(
  proj_obj = proj_obj2,
  point_param = c(1.5, 0.2), # size, alpha, color
  plot_limits = c(-1, 0.75), 
  title = "b3", 
  cex = 2, 
  axis_text_size = 5,
  is_color = TRUE) + scale_color_manual(values = c("#8dd3c7", "#fdb462", "#bebada"))
```

@@

free(trimesh_removed_pbmc) + free(trimesh_removed_pbmc_best) +
  pbmc_proj_umap_model1 + pbmc_proj_tsne_model1_best +
  pbmc_proj_umap_model2 + pbmc_proj_tsne_model2_best +
  plot_layout(nrow=2, byrow=FALSE) &
  theme(legend.position='none')
```
